Inferring the distribution and rate of Earth-like exoplanets \\ from noisy individual-planet inferences

### 2014-03-28 document/ms.tex
text stuff

\begin{abstract}
Although no true extra-solar Earth analog is known, hundreds of planets have
been found around Sun-like stars that are either Earth-sized but on shorter
periods, or else on year-long orbits but somewhat bigger than Earth.
These populations ought to permit a (possibly highly uncertain) extrapolated
measurement of the probability or rate at which Sun-like stars host true Earth
analogs.
Here we build a set of probabilistic models of the population of known small
planets around G dwarfs.
The models all take into account the transit probability and the completeness
(or detection efficiency) as a function of planet radius and orbital period.
Because we forward model the data, we do not ever have to perform the
traditional weighting of the data by the inverse detection efficiency.
Our most sophisticated models do not assume that the joint radius--period
distribution is separable, and correctly take into account the significant
observational uncertainties on the planet radii, using the machinery of
hierarchical inference.
We reconsider and confirm with weaker assumptions prior results on the
distribution of exoplanet radii that show a flattening of the distribution
below a few Earth radii.
One output of our models is fully marginalized estimates---marginalizing out
observational uncertainties and all distribution parameters---of the rate
density of true Earth analogs, expressed as a number of planets per star per
natural logarithmic interval of period and radius, evaluated at the properties
of Earth.
The rate density we infer is an extrapolation; it depends strongly on the
permitted flexibility of the distribution model, but it tends to come out
at around ten percent, with large uncertainty.
\end{abstract}

@@

This study is novel in a number of ways.
We create probabilistic information about the full exoplanet population
\emph{not} by weighting the detected objects by the inverse detection
efficiency (\citealt{dong, petigura}), but rather by forward modeling the
observed distribution subject to the detection efficiency.
\hoggtodo{Check that Dong \& Zhu do indeed weight by inverse probability.}
Re-weighting the data by inverse selection probability---a method called
V-max in the quasar and galaxy luminosity function contexts---is not wrong,
but it produces a higher-variance estimate of the population than a justified
likelihood approach.
One simple intuition along these lines is that the inverse probability weights
can be large in dynamic range, so the effective number of samples, after
weighting, can be much smaller than the actual number of samples.
Another is that it tends to upweight most strongly the most marginal
detections in the data sample, since these tend to be the samples that lie in
the region of lowest completeness.  In this work, we build a justified
likelihood of the observed data, taking into account the completeness function
or detection efficiency.
Our likelihood function is a variable-rate Poisson likelihood; it makes the
fundamental assumption that the data points are independently drawn from the
model, but is agnostic about all other properties of the data or distribution
functions.
Indeed, the assumptions behind this study are weaker than those of any
previous study of the period and radius distribution of the Earth-analog
population (\citealt{dong, petigura}).

>>>>

\begin{abstract}
Although no true extra-solar Earth analog is known, hundreds of planets have
been found around Sun-like stars that are either Earth-sized but on shorter
periods, or else on year-long orbits but somewhat bigger than Earth.
These populations ought to permit a (possibly highly uncertain) extrapolated
measurement of the probability or rate at which Sun-like stars host true Earth
analogs.
Here we build a set of probabilistic models of the population of known small
planets around G dwarfs.
The models take into account the transit probability and the completeness
(or detection efficiency) as a function of planet radius and orbital period.
Because we forward model the data, we do not ever have to perform any kind of
inverse-detection-efficiency weighting of the data.
Our most sophisticated models do not assume that the joint radius--period
distribution is separable.
They also correctly take into account the significant observational
uncertainties on the planet radii, using the machinery of hierarchical
inference.
We reconsider and confirm with weaker assumptions prior results on the
distribution of exoplanet radii that show a flattening of the distribution
below a few Earth radii.
One output of our models is fully marginalized estimates---marginalizing out
observational uncertainties and all distribution parameters---of the rate
density of true Earth analogs, expressed as a number of planets per star per
natural logarithmic interval of period and radius, evaluated at the properties
of Earth.
The rate density we infer is an extrapolation; it depends strongly on the
permitted flexibility of the distribution model, but it tends to come out
at around ten percent, with large uncertainty.
\end{abstract}

@@

This study is novel in a number of ways.
We create probabilistic information about the full exoplanet population
\emph{not} by weighting the detected objects by the inverse detection
efficiency (\citealt{petigura}), but rather by forward modeling the
observed distribution subject to the detection efficiency
(\dfmtodo{CITE TREMAINE?} \citealt{dong}).
Re-weighting the data by inverse selection probability---a method called
V-max in the quasar and galaxy luminosity function contexts---is not wrong,
but it produces a higher-variance estimate of the population than a justified
likelihood approach; it is higher variance because
the effective number of samples, after
weighting, can be much smaller than the actual number of samples.
In this work, we build a justified
likelihood of the observed data, taking into account the completeness function
or detection efficiency.
Our likelihood function is a variable-rate Poisson likelihood; it makes the
fundamental assumption that the data points are independently drawn from the
model, but is agnostic about all other properties of the data or distribution
functions.
Indeed, the assumptions behind this study here are weaker than those of any
previous study of the period and radius distribution of the Earth-sized
exoplanet population (\dfmtodo{CITE TREMAINE?} \citealt{dong, petigura}).
